% !TEX root =  ipdps18.tex

\section{Lower Bound}
\label{sec:lowerbound}
% Primary: Yves (does that go into a subsec of the algorithms or models?)

We now derive a lower bound for optimal platform waste.  When we assess the
performance of the scheduling algorithms presented in \Cref{sec:algorithms}, we
also compare their relative performance to this lower bound (in
\Cref{sec:results}).

We envision a (theoretical) scenario in which the platform operates in
steady-state, a constant number of jobs per application class spanning the
entire platform.  We also assume that the I/O bandwidth $\bandavail$ available
for CR operations remains constant throughout execution. This amounts to
ignoring initial input and final output I/O operations, or more precisely, to
assuming these operations span the entire execution of the jobs.  Without this
assumption, we would need to account for job durations; this renders the
steady-state analysis intractable.  Given above, we determine the optimal
checkpointing period for each application class with the objective to minimize
the total waste of the platform; or equivalently, to maximize the total
throughput of the platform. To complicate this analysis, these optimal periods
may not be achievable, hence we derive a lower bound of the optimal waste.

In steady-state operation, there are $\nbapp{i}$ jobs of class $\app{i}$, each
using $\nbnodes{i}$ nodes, and with checkpoint time $\ckpt{i}$. Because we
orchestrate checkpoints to avoid CR-CR interferences, we have $\ckpt{i} =
\frac{\size{i}}{\bandavail}$, where $\size{i}$ denote the size of the
checkpoint file of all jobs of class $\app{i}$.  The waste of a job is the
ratio of time the job spends doing resilience operations by the time it does
useful work. The time spent performing resilience operations include the time spent
during each period to checkpoint; and in case of failure, the time to rollback
to the previous checkpoint and the time to recompute lost work.
%We assume that the recovery time $\reco{i}$ is equivalent to the checkpoint
%time  $\ckpt{i}$.
We can express the waste $\wasteapp{i}$ of a job $J_{i}$ of class $\app{i}$
that checkpoints with period $\period{i}$ as follows~\cite{springer-monograph}:

\begin{equation}
\wasteapp{i} = \wastefct{i}{\ckpt{i}} = \frac{\ckpt{i}}{\period{i}} +
\frac{\nbnodes{i}}{\mtbfplat}(\frac{\period{i}}{2} + \reco{i})
\label{eq.wasteAi}
\end{equation}

Let $\wasteplat$ be the waste of the platform. We define this as the
weighted arithmetic mean of the $\wasteapp{i}$ for all applications,
where each application is weighted by the number of computing nodes
it uses:

\begin{equation}
\wasteplat = \sum_i \frac{\nbapp{i} \nbnodes{i}}{\nbnodesplat} \wasteapp{i}
\label{eq.waste}
\end{equation}

In the absence of I/O constraints, the checkpointing period can be minimized
for each job independently. Indeed, the optimal period for a job
of class $\app{i}$ is obtained by minimizing $\wasteapp{i}$ in 
Equation~\eqref{eq.wasteAi}.

Differentiating and solving
$$\frac{\delta \wasteapp{i}}{\delta \period{i}} = - \frac{\ckpt{i}}{\period{i}^{2}} + \frac{\nbnodes{i}}{2 \mtbfplat} = 0$$
we readily derive that
\begin{equation}
\period{i} = \sqrt{2 \frac{\mtbfplat}{\nbnodes{i}} \ckpt{i}} = \sqrt{2 \mu_{i} \ckpt{i}}
\label{eq.daly}
\end{equation}
where $\mu_{i}$ is the MTBF of  class $\app{i}$ applications, and we retrieve the Daly period
$\period{i} = \period{Daly}(J_{i})$.

However, I/O constraints may impose the use of sub-optimal periods. If each job
of class $\app{i}$ checkpoints in time $\ckpt{i}$ during its period $\period{i}$ (hence
without any contention), it uses the I/O device during a fraction $\frac{\ckpt{i}}{\period{i}}$ of the time.
The total usage fraction of the  I/O device is $\ioconstraint = \sum_{i} \frac{\nbapp{i} \ckpt{i}}{\period{i}}$
and cannot exceed $1$. Therefore, we have to solve the following optimization problem: find
the set of values $\period{i}$ that minimize $\wasteplat$ in Equation~\eqref{eq.waste} subject to the I/O constraint:

\begin{equation}
\ioconstraint = \sum_{i} \frac{\nbapp{i} \ckpt{i}}{\period{i}} \leq 1
\label{eq.IOconstraint}
\end{equation}

Hence the optimization problem is to minimize:
\begin{equation}
\wasteplat = \sum_i \frac{\nbapp{i} \nbnodes{i}}{\nbnodesplat}  \left( \frac{\ckpt{i}}{\period{i}} +
\frac{\nbnodes{i}}{\mtbfplat}(\frac{\period{i}}{2} + \reco{i}) \right)
\label{eq.totalwaste}
\end{equation}
subject to Equation~\eqref{eq.IOconstraint}.
Using the Karush-Kuhn-Tucker conditions~\cite{Boyd2004}, we know that there exists a nonnegative constant
$\lambda$
such that
$$- \frac{\delta \wasteplat}{\delta \period{i}} = \lambda \frac{\delta \ioconstraint}{\delta \period{i}}$$
for all $i$. We derive that
$$\frac{\nbapp{i} \nbnodes{i} \ckpt{i}}{\nbnodesplat \period{i}^{2}} -    \frac{\nbapp{i} \nbnodes{i}^{2}}{2 \mtbfplat \nbnodesplat} = - \lambda \frac{\nbapp{i} \ckpt{i}}{\period{i}^{2}}
$$
for all $i$. This leads to:
 \begin{equation}
\period{i} = \sqrt{\frac{2 \mtbfplat  \nbnodesplat}{\nbnodes{i}^{2}} \left(\frac{\nbnodes{i}}{\nbnodesplat} +\lambda \right) \ckpt{i}}
  \label{eq.KKT}
\end{equation}
for all $i$. Note that when $\lambda=0$, Equation~\eqref{eq.KKT} reduces to Equation~\eqref{eq.daly}.

Because of the I/O constraint in Equation~\eqref{eq.IOconstraint}, we choose
for $\lambda$ the minimum value such that Equation~\eqref{eq.IOconstraint} is
satisfied. If $\lambda \neq 0$, this will lead to periods $P_{i}$ larger than
the optimal value of Equation~\eqref{eq.daly}. Note that there is no
closed-form expression for the minimum value of $\lambda$, it has to be found
numerically.

Altogether, we state our main result:

\begin{theorem}
     In the presence of I/O constraints, the optimal checkpoint periods are given by
     Equation~\eqref{eq.KKT}, where $\lambda$ is the smallest non-negative value such
     that Equation~\eqref{eq.IOconstraint} holds. The total platform waste is then
     given by Equation~\eqref{eq.totalwaste}.
\end{theorem}

The optimal periods may not be achievable, because
Equation~\eqref{eq.IOconstraint} is a necessary, but not sufficient condition.
Even though the total I/O bandwidth is not exceeded, meaning there is enough
capacity to take all the checkpoints at the given periods, we would still need
to orchestrate these checkpoints into an appropriate, periodic, repeating
pattern.  In other words, we only have a lower bound of the optimal platform
waste.
